Response to reviewers

### 2022-02-07 paper/paper.Rmd
Modified Section 2.2. as per Reviewer 2's comment.

```{r theme, eval = FALSE}
theme_set(theme_bw(base_size = 18) +
            theme(plot.background = element_rect(fill = 'transparent', colour = NA), axis.line.y = element_line(color = "black", linetype = "solid"),
                  plot.title.position = "plot",
                  plot.title = element_text(size = 24),
                  panel.background  = element_rect(fill = 'transparent', colour = NA),
                  legend.background = element_rect(fill = 'transparent', colour = NA),
                  legend.key        = element_rect(fill = 'transparent', colour = NA)
                  ) )
```

@@

<!--AE comment: What does “yearly mean hourly wages with years of workforce experience” mean? Is that one variable or two?-->
<!--AE comment: Was your goal to collect data from 1994-2018 or 1979-2018? Page 4 states 1994 onward, but from the figures later it appears that you collected data starting in 1979 (which makes more sense to me).-->
<!--AE comment: “The NLSY79 data used in Singer and Willett (2003) contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994.” Might work better split into two sentences. -->
The NLSY79 data used in @SingerJudithD2003Alda contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994. In addition, the cohort is restricted to male high-school dropouts who first participated in the study at age 14-17 years. 
<!--AE comment: Thus the target of our study is to collect the same variables for the same time period and extending through to 2018, the most recent year reported.-->Thus the target data set is to collect the same variables for the extended time frame of 1994 through to 2018, the most recent year reported.

>>>>

```{r theme, eval = FALSE}
theme_set(theme_bw(base_size = 18) +
            theme(plot.background = element_rect(fill = 'transparent', colour = NA), axis.line.y = element_line(color = "black", linetype = "solid"),
                  plot.title.position = "plot",
                  plot.title = element_text(size = 24),
                  panel.background  = element_rect(fill = 'transparent', colour = NA),
                  legend.background = element_rect(fill = 'transparent', colour = NA),
                  legend.key        = element_rect(fill = 'transparent', colour = NA)
                  ))
```

@@

The NLSY79 data used in @SingerJudithD2003Alda contains the longitudinal measurements of the male high school dropouts who first participated in the study at age 14-17 years from 1979 through to 1994. This dataset comprises of several variables as follows:

1. ID: the respondents' ID. 
2. EXPER: temporal scale, i.e., the length of time (years) in the workforce, starting on the respondents' first day at work.
3. LNW: natural logarithm of wages, adjusted with 1990's inflation rate. 
4. BLACK: binary variables, the respondents are coded to 1 because they were Black and 0 otherwise. 
5. HGC: the highest grade completed.
6. UERATE: unemployment rate of the year of survey. When missing, the variable is set to be 7.875 (the average rate).

From the database, we will refresh this data with a more up to date period, which is from survey year 1979 to survey year 2018 (the most recent year published). We will  also modify the variables (for example, we create a race variable, instead of the binary variable "Black") and add some variables so that more examples of data exploration can be displayed in teaching. The variables added are year of survey, age of individual in 1979, year when the highest grade completed, number of jobs that the individual had in the corresponding year of survey, total number of hours the individual usually works per week, the year when individual starting to work, and the number of years the individual worked.

In addition to the dropouts data, we will also make the up to high school graduates' wages available, including the females, and the demographic data of the NLSY79 cohort available in separate datasets. Thus, we will create three datasets as follows:

1. The demographic data of the NLSY79 cohort.
2. The wages data of the up to high school graduates.
3. The wages data of the high school dropouts, which is the refreshed version of @SingerJudithD2003Alda's data.

### 2022-02-07 paper/paper.tex
Edited the paper and add the response.
Modified Section 2.2. as per Reviewer 2's comment.

The NLSY79 data used in Singer and Willett (2003) contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994. In addition, the cohort is restricted to male high-school dropouts who first participated in the study at age 14-17 years.
Thus the target data set is to collect the same variables for the extended time frame of 1994 through to 2018, the most recent year reported.

@@

According to Wickham (2014), tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table. The raw data does not comply with these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables. This is done using the \texttt{tidyverse} suite of packages (Wickham, Averick, et al. 2019), \texttt{tidyr} (Wickham 2020) to pivot the data into long form, \texttt{dplyr} (Wickham, François, et al. 2020), and \texttt{stringr} (Wickham 2019) for creating new variables, and levels of factors by text wrangling. The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data. The code for tidying the data are demonstrated at \url{CENSORED/articles/raw-to-input-data.html} but also described in the subsequent subsections.

@@

The number of jobs (\texttt{number\_of\_jobs}) for each respondent per year is computed from the number of non-missing values of hourly wage \emph{and} hours worked. If either the hourly wage or hours worked is missing, we do not tally this.

@@

As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979. Hence, we validate whether all of the respondents were in this range. Additionally, the \href{https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction}{NLSY} also provides the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic). To validate this, we used the \texttt{demog\_nlsy79}, i.e., the data with the survey years 1979 sample.
Tables \ref{tab:age-table} and \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.

@@

\caption{\label{tab:age-table}The frequency table of the age at the start of the survey (?CHECK) in the full NSLY79 data}
\centering
\begin{tabular}[t]{rr}
\toprule
Age & Number of individuals\\
\midrule
15 & 1,265\\
16 & 1,550\\
17 & 1,600\\
18 & 1,530\\
19 & 1,662\\
20 & 1,722\\
21 & 1,677\\
22 & 1,680\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{center}\includegraphics[width=468px]{figures/compare_xp_to_yrworkforce-1} \end{center}

>>>>

The NLSY79 data used in Singer and Willett (2003) contains the longitudinal measurements of the male high school dropouts who first participated in the study at age 14-17 years from 1979 through to 1994. This dataset comprises of several variables as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  ID: the respondents' ID.
\item
  EXPER: temporal scale, i.e., the length of time (years) in the workforce, starting on the respondents' first day at work.
\item
  LNW: natural logarithm of wages, adjusted with 1990's inflation rate.
\item
  BLACK: binary variables, the respondents are coded to 1 because they were Black and 0 otherwise.
\item
  HGC: the highest grade completed.
\item
  UERATE: unemployment rate of the year of survey. When missing, the variable is set to be 7.875 (the average rate).
\end{enumerate}

From the database, we will refresh this data with a more up to date period, which is from survey year 1979 to survey year 2018 (the most recent year published). We will also modify the variables (for example, we create a race variable, instead of the binary variable ``Black'') and add some variables so that more examples of data exploration can be displayed in teaching. The variables added are year of survey, age of individual in 1979, year when the highest grade completed, number of jobs that the individual had in the corresponding year of survey, total number of hours the individual usually works per week, the year when individual starting to work, and the number of years the individual worked.

In addition to the dropouts data, we will also make the up to high school graduates' wages available, including the females, and the demographic data of the NLSY79 cohort available in separate datasets. Thus, we will create three datasets as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The demographic data of the NLSY79 cohort.
\item
  The wages data of the up to high school graduates.
\item
  The wages data of the high school dropouts, which is the refreshed version of Singer and Willett (2003)'s data.
\end{enumerate}

@@

According to Wickham (2014), tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table. The raw data does not comply with these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables.This is done using the \texttt{tidyverse} suite of packages (Wickham, Averick, et al. 2019), i.e., \texttt{tidyr} (Wickham 2020) to pivot the data into long form, \texttt{dplyr} (Wickham, François, et al. 2020) and \texttt{stringr} (Wickham 2019) to create new variables from the downloaded data from the database, and create levels of factors by text wrangling. The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data. The code for tidying the data are demonstrated at \url{CENSORED/articles/raw-to-input-data.html} but also described in the subsequent subsections.

@@

The number of jobs (\texttt{number\_of\_jobs}) for each respondent per year is computed from the number of non-missing values of hourly wage. Even if the information of hours worked exists for a particular observation, we do not tally when the hourly wage is missing.

@@

As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979. Hence, we validate whether all of the respondents were in this range in the data we extracted. Additionally, the \href{https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction}{NLSY} also provides the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic). To validate this, we used the \texttt{demog\_nlsy79}, i.e., the data with the survey years 1979 sample. Tables \ref{tab:age-table} and \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.

@@

\caption{\label{tab:age-table}The frequency table of the age at the start of the survey in NSLY79 cohort}
\centering
\begin{tabular}[t]{rr}
\toprule
Age & Number of individuals\\
\midrule
15 & 1,265\\
16 & 1,550\\
17 & 1,600\\
18 & 1,530\\
19 & 1,662\\
20 & 1,722\\
21 & 1,677\\
22 & 1,680\\
\bottomrule
\end{tabular}
\end{table}

@@

### 2022-02-07 response/response1.Rmd
Modified Section 2.2. as per Reviewer 2's comment.

>>>>

\begin{description}
\item[Reviewer 2:]\textcolor{violet}{I would like more discussion about the textbook selected as the exemplar. It sounds like this dataset is used in many books, why choose that one? The text should include the name of the book, so readers do not need to flip to the citations in order to learn what it was. From the name, I am guessing that the book focuses heavily on this dataset, which is why it makes sense to reproduce their subset. Be explicit about this.}
\end{description}

\begin{description}
\item[Reviewer 2:]\textcolor{violet}{When listing the files that come zipped from the data source, be more explicit about the file formats. Comma separated values instead of csv, etc.}
\end{description}

* We have made it explicit.

\begin{description}
\item[Reviewer 2:]\textcolor{violet}{Before the data cleaning section I would love a glimpse into the future to see what the target dataset would look like. What are the observations? What are the variables? This would allow me to follow the narrative of data wrangling more easily. Perhaps this could be put into section 2.2, which is titled Target data. Then, when starting to describe the raw data in 3,1, explain the same pieces and explicitly explain why the raw data it is not tidy to begin with.}
\end{description}

\begin{description}
\item[Reviewer 2:]\textcolor{violet}{Before the data cleaning section I would love a glimpse into the future to see what the target dataset would look like. What are the observations? What are the variables? This would allow me to follow the narrative of data wrangling more easily. Perhaps this could be put into section 2.2, which is titled Target data. Then, when starting to describe the raw data in 3.1, explain the same pieces and explicitly explain why the raw data it is not tidy to begin with.}
\end{description}

* We have added the explanation of the observations and the variable in Section 2.2.
